Empirical normative theories

### 2020-07-16 simple.tex
Removed a bit more stuff; changed some commands

The object of study in a given decision situation is the deliberated preferences of a given individual $i$. 
A decision situation concerns a topic $\alltopic$, given a set of arguments $\allargs$, given a certain query protocol, and over a certain time frame. The relation $\ileadstost$, to be defined shortly, models the reaction of $i$ to arguments about the topic. The decision situation will be defined as a triple $(\alltopic, \allargs, \ileadstost)$ satisfying some conditions to be defined after having presented those three fundamental elements.

The decision situation relates to a topic denoted by $\alltopic$, which is a set of propositions $\prop \in \alltopic$ about which we are interested of knowing the deliberated preferences of $i$. Propositions are not described further, but are supposedly understandable by $i$ (for example, they could be sentences in some natural language). Alternatively, in the case of an analyst helping $i$ to make a decision, $\alltopic$ is a set of propositions about which $i$ is interested to know his own deliberated preferences. I assume $\alltopic$ is closed under negation and introduce a symbol $¬$ for negating a proposition: if $\prop \in \alltopic$, then $¬\prop \in \alltopic$, and $¬(¬\prop) = \prop$.

With the decision situation comes also a set of all arguments $\allargs$. This set contains all the arguments that may possibly be considered relevant by $i$ for his decision problem, and possibly more.
Importantly, the notion of deliberated preferences does not require to constrain a priori $\allargs$ to some set of arguments that would fit some precise notion of relevancy, coherence, or even well-formedness. This permits to avoid introducing inadequate normative principles into the deliberated preferences: only the individual $i$ is then considered legitimate to dictate what is a relevant argument. Under that view, $\allargs$ may contain anything that can possibly be considered as an argument by anyone, under the widest possible conception of an argument. The notion of deliberated judgment can however also be applied when considering a restricted set of arguments, for example, the arguments that have been put forward by some specific set of experts when talking about the topic $\alltopic$.

@@

\begin{example}
	Consider a decision problem where a set of alternatives $\allalts$ is given, containing food products among which $i$ would like to form a deliberated preference (taking into account the effects of food on health, price, pleasure, morality issues related to the production process, and so on). The topic could be defined as $\alltopic$ = $\set{\prop_\alt, ¬\prop_\alt, \forall \alt \in \allalts}$, where $\prop_\alt$ is the proposition according to which $\alt$ is one of the best alternatives among $\allalts$ (there is no strictly better alternatives among $\allalts$, from $i$’s point of view), and $¬\prop_\alt$ is the proposition according to which $\alt$ is not one of the best alternatives among $\allalts$ (some $\alt' \in \allalts$ is a strictly better alternative, from $i$’s point of view). 
	The set $\allargs$ is defined as the set of all strings, because in this example it is considered acceptable to restrict arguments to those that can take a textual form.
\end{example}

Define $\Phi = \alltopic × \set{\text{possible}, \text{sure}}$. Elements in $\Phi$ will also be called propositions, which should create no ambiguity. 
The relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$, pronounced “always leads to”, has the following semantics.
Whenever $(\ar, \ar') \ileadstost (\prop, \text{sure})$, $i$ always considers that $\prop$ is sure (meaning that $¬\prop$ is excluded), when presented with these two arguments, and whenever $(\ar, \ar') \ileadstost (\prop, \text{possible})$, $i$ always considers that $\prop$ is “at least” possible (meaning that $¬\prop$ may be possible as well), when presented with these two arguments. In both cases, the term “always” means that the consideration is stable over time, including after having presented other arguments to $i$. 
Two query protocols presented below will illustrate how these semantics might be satisfied. 
However, I voluntarily leave the querying protocol not precisely specified in general, as multiple reasonable choices are possible. The results of this article depend only on assumptions about $\ileadstost$.

@@

\begin{definition}[Decision situation]
	A decision situation is a topic $\alltopic$ closed under negation, a set of arguments $\allargs$, an “always leads to” relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$ satisfying (A1) and (A2) and using unordered pairs of arguments, $\Phi$ being the set of propositions $\alltopic × \set{\text{sure}, \text{possible}}$.
\end{definition}

@@

It is also necessary to capture the evolution of the position of $i$ towards arguments over time: a repeated query using a given pair of arguments may yield different answers, for example because the repeated question has been interleaved with another query containing other arguments (which $i$ may still have in mind when answering the second repetition), or for any other known or unknown reason. The important information for us is the \emph{set} of answers that $i$ could give to a query involving two given arguments and a given topic, among the set $\{\prop, ¬\prop, \text{both}\}$. The set designates all possible answers over time and over different order of asking the queries.
Hence, given $\prop \in \alltopic$, define a relation $\ileadsto^\prop \subseteq (\allargs × \allargs) × \powersetz{\{\prop, ¬\prop, \text{both}\}}$, where $\powersetz{B}$ designates the subsets of $B$ excluding the emptyset, with the semantics that $(\ar, \ar') \ileadsto^\prop C$, with $C \subseteq \set{\prop, ¬\prop, \text{both}}$, iff $\forall c \in C$, it is observable at least once that $i$, when presented with $(\ar, \ar')$, considers $c$ valid in her current state of mind. The querying protocol is supposedly made so that for any $\prop \in \alltopic$, $\ileadsto^\prop = \ileadsto^{¬\prop}$.

@@

Define $\ileadstost$ from $\ileadsto^\prop$ as follows, for any $\prop \in \alltopic$: $(\ar, \ar') \ileadstost (\prop, \text{possible}) ⇔ (\ar, \ar') \ileadsto^\prop C$ with $C \subseteq \set{\prop, \text{both}}$; and $(\ar, \ar') \ileadstost (\prop, \text{sure}) ⇔ (\ar, \ar') \ileadsto^\prop \set{\prop}$.

@@

\subsection{Deliberated preference}
The following important definition is required in order to define the deliberated preference of $i$. An argument $\ar$ is decisive for $\phi$, in short, $\phi\text{-dec}$ iff $\forall \ar' \in \allargs: (\ar', \ar) \ileadstost \phi$. 
\footnote{Equivalently: $\ar \ \phidec ⇔ [\ileadstostinv(\phi)](\ar) = \allargs$, and $\exists \ar \ \phidec ⇔ [\ileadstostinv(\phi)]^{-1}(\allargs) ≠ \emptyset$.}
I also say that $\ar$ is $\propsuredec$, $\proppossdec$, or $¬\proppossdec$, when $\phi=(\prop, \text{sure})$, $\phi=(\prop, \text{possible})$, or $\phi=(¬\prop, \text{possible})$, respectively.
\footnote{$\propsuredec ⇔ \forall \ar': (\ar', \ar) \ileadstosst \prop ⇔ \forall \ar': ¬[(\ar', \ar) \ileadstope ¬\prop]$, $\proppossdec ⇔ \forall \ar': (\ar', \ar) \ileadstopst \prop$. Note that “both” ≠ “we donno”: in case “both”, we know how to prove that both are reasonable options, in the sense that neither is certain. We have excluded some possibilities (and some other model could disagree with the “both” conclusion).}

\begin{proposition}
	\label{thm:nocontrdec}
	For any decision situation, a) $\exists \ar \ \phidec ⇒ \nexists \ar' \ ¬\phidec$ and b) $\exists \ar \ \propsuredec ⇒ \nexists \ar' \ ¬\propsuredec$.
\end{proposition}
\begin{proof}
	About a), when $\ar$ is $\phidec$, given any $\ar'$, $(\ar, \ar') \ileadstost \phi$ hence $(\ar, \ar') \ileadstoe \phi$ (using (A1)), hence $(\ar, \ar') \nileadstost ¬\phi$. b) obtains similarly using (A2).
\end{proof}

The deliberated preference of $i$ maps each pair $(\prop, ¬\prop)$ to one among four possibilities, depending on the existence of decisive arguments supporting the propositions about $\prop$. The mapping is defined as follows.
\begin{description}
	\item[$\prop$ is sure] iff $\exists \ar \ \propsuredec$
	\item[$¬\prop$ is sure] iff $\exists \ar \ ¬\propsuredec$
	\item[both $\prop$ and $¬\prop$ are possible] iff $\exists \ar \ \proppossdec$ and $\exists \ar \ ¬\proppossdec$
	\item[$(\prop, ¬\prop)$ is undecided] iff neither $\prop$ is sure, nor $¬\prop$ is sure, nor both are possible.
\end{description}
\Cref{thm:nocontrdec} shows that the cases are separate.
\footnote{Example of $(\prop, ¬\prop)\text{-undecided}$: $\forall (\ar, \ar'), (\ar, \ar') \ileadstoe (¬\prop, \text{sure}), (\ar, \ar') \ileadstoe (\prop, \text{sure})$.}
\footnote{Example of $\overline{¬\prop\text{-CC}} ∧ \prop \in \iposs$ (thus with $\prop\text{-CC}$ and no $(\prop, ¬\prop)$-unstability): $\forall (\ar, \ar'), (\ar, \ar') \ileadstose \prop, (\ar, \ar') \ileadstope ¬\prop, (\ar, \ar') \ileadstopst \prop$. In such a situation, it is unclear whether $\prop$ holds rather than $¬\prop$ or both $\prop$ and $¬\prop$.}
\footnote{Something like Axiom JU should be sufficient for CC. We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}

It should be clear that the validity of the conclusions obtained from any theory of deliberated preference rests on the acceptance of this just given normative definition, namely, it is to be considered meaningful that the deliberated preference of $i$ be defined in such a way, given the query protocol used for defining $\ileadstost$ in a particular application.

\commentOC{TODO}
We want to leave “undecided” as an option: the model takes no position, does not know about $(t, ¬t)$. But we want to exclude $t$ possible and $¬t$ unknown. Thus $t$ possible should imply $¬t$ possible or $t$ sure. If this holds, we can simplify the status of the propositions: only determine whether $t$ is possible for all $t \in \alltopic$, and from there, we know everything. This result can be obtained with: assume $s$ decisive for $t$ poss and no $s$ is decisive for $¬t$ poss, then, for some $(s_1, s)$ leading to $t$ sure and $(s_2, s)$ leading to $\set{t, ¬t}$, it is possible to form $(s_2, s_1, s)$, and if this leads to $t$ sure, it should imply that $s + s_1$ is decisive for $t$ sure; and if this leads to $¬t$ possible, then this should imply that $s+s_2$ is decisive for $¬t$ possible.

This information is summarized by defining a subset $\iPhi \subseteq \Phi$ of propositions that represent $i$’s deliberated preference, defined as follows.
\begin{description}
	\item[$\prop$ is sure] ⇒ $(\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \in \iPhi ∧ (¬\prop, \text{possible}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$
	\item[both $\prop$ and $¬\prop$ are possible] ⇒ $(\prop, \text{possible}) \in \iPhi ∧ (¬\prop, \text{possible}) \in \iPhi ∧ (\prop, \text{sure}) \notin \iPhi ∧ (¬\prop, \text{sure}) \notin \iPhi$
	\item[$(\prop, ¬\prop)$ is undecided] ⇒ $\phi \notin \iPhi$ for all $\phi$ about $\prop$.
\end{description}
Because this defines unequivocally the membership of every proposition about $\prop$ for every $\prop \in \alltopic$, this suffices to define $\iPhi$; and it follows that the reverse implications hold as well.

The following propositions can help get convinced that the proposed definition of deliberated preference makes sense.
\begin{proposition}
	a) $\phi \in \iPhi ⇒$ some $\ar$ is $\phidec$. b) $(\prop, \text{possible}) \in \iPhi$ iff $\exists \ar \ \proppossdec ∧ (\exists \ar \ \propsuredec ∨ \exists \ar \ ¬\proppossdec)$. c) $(\prop, \text{sure}) \in \iPhi ⇒ (\prop, \text{possible}) \in \iPhi$. d) $\phi \in \iPhi ⇒ ¬\phi \notin \iPhi$. e) $(\prop, ¬\prop)$ undecided $⇔ \phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$. f) If $\phi \in \iPhi$ for some $\phi$ about $\prop$, then $\phi \in \iPhi ⇔ ¬\phi \notin \iPhi$ for all $\phi$ about $\prop$. g) If $\ileadstost$ does not distinguish sure from possible, no pair of proposition is mapped to “both are possible”.
\end{proposition}
\begin{proof}
	a) is a consequence of A2. b) and c) follow from the definition of $\iPhi$. d) is obtained using a), \cref{thm:nocontrdec}, and the contrapositive of a). About e), we must show that $\phi \notin \iPhi ∧ ¬\phi \notin \iPhi$ for some $\phi$ about $\prop$ implies $\phi' \notin \iPhi$ for all $\phi'$ about $\prop$. Assume $\phi=(l, \text{possible})$ and $¬\phi = (¬l, \text{sure})$ are not in $\iPhi$, with $l = t$ or $l = ¬t$. We need to prove that $(l, \text{sure})$ and $(¬l, \text{possible})$ are not in $\iPhi$ either. We obtain $(l, \text{sure}) \notin \iPhi$ using the contrapositive of c). And $(¬l, \text{possible}) \notin \iPhi$ because $(¬l, \text{possible}) \in \iPhi$ would require another $\phi'$ about $\prop$ to be present in $\iPhi$, as can be checked from the definition of $\iPhi$. f) follows from d) together with the fact that $(\prop, ¬\prop)$ is (by definition) not undecided when $\phi \in \iPhi$ for some $\phi$ about $\prop$, hence, from e), $\phi \in \iPhi ∨ ¬\phi \in \iPhi$. Finally, g) is because in that case, $\exists \ar \ \proppossdec ⇔ \exists \ar \ \propsuredec$, thus using \cref{thm:nocontrdec}, $\exists \ar \ \proppossdec ⇒ \nexists \ar' \ ¬\proppossdec$.
\end{proof}

@@

\section{Empirical theories}
\subsection{Simple to complex}
From an additive single-answer empirical theory, we can define a function: $\forall \ar_1 \gbeats \ar_0: f(\ar_1, \ar_0) = \ar_0 + {\gbeatsinv}(\ar_1) = \set{\ar_0 + \ard_2 \suchthat \ard_2 \in {\gbeatsinv}(\ar_1)}$, and $f(\ar_1, \ar_0) = \emptyset$ otherwise. Thus, $f = \set{((\ar_1, \ar_0), \ar_2) \suchthat \exists \ard_2 \in \allargs \suchthat \ard_2 \gbeats \ar_1 \gbeats \ar_0 \land \ar_2 = \ar_0 + \ard_2}$.
Also, from the previous definition, $g(\ar_0)(\ar_0 + \argsd_2) = dom(f^{-1}(\ar_0 + \argsd_2) \cap (\allargs × \set{\ar_0})) = {\gbeats}(\argsd_2) \cap {\gbeatsinv}(\ar_0)$ because $(\ar_1, \ar_0) \in f^{-1}(\ar_0 + \argsd_2) ⇔ \exists \ar_2 \in \ar_0 + \argsd_2 \suchthat \ar_2 \in f(\ar_1, \ar_0) ⇔ \ar_1 \gbeats \ar_0 \land \exists \ard \in \argsd_2 \suchthat \ard \gbeats \ar_1$.
\commentOC{Could be false. Define }

@@

Given any $(\alltopic, ¬, \allargs, I)$, an \emph{observation protocol} is a binary tree whose non-leaf nodes are among $I × \allhist × \Phi$ and whose two edges out of any non leaf node are labeled by “yes” and “no”; the leaf nodes being simply $\emptyset$. The observation protocol is said to assume memory iff for each non-leaf node $n = (i, \hist, \phi)$ about $i$, any node $n' = (i, {\hist}', \phi')$ about $i$ that is further down the tree continues the history, that is, the sequence ${\hist}'$ starts with $\hist$. Given an observation protocol and observables $(\ileadstoe[i])_{i \in I}$, the \emph{observed path} refers to the path in the tree defined recursively as consisting of the root node and, for each non-leaf node $n = (i, \hist, \phi)$ in the path, the child node along the branch “yes” if $\hist \ileadstoe[i] \phi$ and the child node along the branch “no” otherwise. The \emph{observations} refer to the set of nodes along the observed path.

Given any $(\alltopic, ¬, \allargs, I)$, a \emph{normative claim} about $(\alltopic, ¬, \allargs, I)$ (or about $\allargs$, when the other defined objects are left implicit) is any first-order logic proposition involving the atoms $[(\hist, \phi) \in {\ileadstoe[i]}], \hist \in \allhist, \phi \in \Phi$ which, together with the axioms, permits to deduce that $\phi \in \iPhi$ for some $i \in I$ and $\phi \in \Phi$. The claim is said to be \emph{about} the given set $\allargs$. An observation protocol, together with observables $(\ileadstoe[i])_{i \in I}$, \emph{falsifies} a normative claim iff the falsity of the claim is deducible from the axioms and the observations. A normative claim is \emph{falsifiable} iff there exists a finite observation protocol such that for any observables $(\ileadstoe[i])_{i \in I} \in (\allhist × \Phi)^I$, if the claim is false given these observables, then it is possible to deduce it from the axioms and the observations given the observables.

@@

\begin{proposition}[Trivial claims are not falsifiable]
	Given any $(\alltopic, ¬, \allargs ≠ \emptyset, I)$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Consider any claim about some $i$ and some $\phi$ and any finite observation protocol. We have to show that for some observables, the claim is false, but it is impossible to observe it (meaning, to deduce it from the axioms and observations).

@@

There is another reason for non falsifiability, which does not involve the problem of finiteness.
\begin{proposition}[Minimal non-trivial claims are not falsifiable]
	Given any $(\alltopic, ¬, \allargs, I), \card{\allargs} ≥ 2$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Define $\ar_1$ as the first element of the sequence $\hist$ of the root node of the observation protocol (if the observation protocol is empty, suffices to define any observables that make the claim false). Define the observables so that $\forall \hist \in \allhist: \ar_2, \hist \ileadsto ¬\phi$. This proves that no argument is decisive, hence $\phi \notin \Box\Phi$, and the claim is false. But the observables may also be defined so as to satisfy the protocol (defined as: when going through the protocol, the observations can’t fail the claim).
\end{proof}

>>>>

The object of study in a given decision situation is the deliberated preferences of a given individual $i$. 
A decision situation concerns a topic $\allprops$, given a set of arguments $\allargs$, given a certain query protocol, and over a certain time frame. The relation $\ileadstost$, to be defined shortly, models the reaction of $i$ to arguments about the topic. The decision situation will be defined as a triple $(\allprops, \allargs, \ileadstost)$ satisfying some conditions to be defined after having presented those three fundamental elements.

The decision situation relates to a topic denoted by $\allprops$, which is a set of propositions $\prop \in \allprops$ about which we are interested of knowing the deliberated preferences of $i$. Propositions are not described further, but are supposedly understandable by $i$ (for example, they could be sentences in some natural language). Alternatively, in the case of an analyst helping $i$ to make a decision, $\allprops$ is a set of propositions about which $i$ is interested to know his own deliberated preferences. I assume $\allprops$ is closed under negation and introduce a symbol $¬$ for negating a proposition: if $\prop \in \allprops$, then $¬\prop \in \allprops$, and $¬(¬\prop) = \prop$.

With the decision situation comes also a set of all arguments $\allargs$. This set contains all the arguments that may possibly be considered relevant by $i$ for his decision problem, and possibly more.
Importantly, the notion of deliberated preferences does not require to constrain a priori $\allargs$ to some set of arguments that would fit some precise notion of relevancy, coherence, or even well-formedness. This permits to avoid introducing inadequate normative principles into the deliberated preferences: only the individual $i$ is then considered legitimate to dictate what is a relevant argument. Under that view, $\allargs$ may contain anything that can possibly be considered as an argument by anyone, under the widest possible conception of an argument. The notion of deliberated judgment can however also be applied when considering a restricted set of arguments, for example, the arguments that have been put forward by some specific set of experts when talking about the topic $\allprops$.

@@

\begin{example}
	Consider a decision problem where a set of alternatives $\allalts$ is given, containing food products among which $i$ would like to form a deliberated preference (taking into account the effects of food on health, price, pleasure, morality issues related to the production process, and so on). The topic could be defined as $\allprops$ = $\set{\prop_\alt, ¬\prop_\alt, \forall \alt \in \allalts}$, where $\prop_\alt$ is the proposition according to which $\alt$ is one of the best alternatives among $\allalts$ (there is no strictly better alternatives among $\allalts$, from $i$’s point of view), and $¬\prop_\alt$ is the proposition according to which $\alt$ is not one of the best alternatives among $\allalts$ (some $\alt' \in \allalts$ is a strictly better alternative, from $i$’s point of view). 
	The set $\allargs$ is defined as the set of all strings, because in this example it is considered acceptable to restrict arguments to those that can take a textual form.
\end{example}

Define $\Phi = \allprops × \set{\text{possible}, \text{sure}}$. Elements in $\Phi$ will also be called propositions, which should create no ambiguity. 
The relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$, pronounced “always leads to”, has the following semantics.
Whenever $(\ar, \ar') \ileadstost (\prop, \text{sure})$, $i$ always considers that $\prop$ is sure (meaning that $¬\prop$ is excluded), when presented with these two arguments, and whenever $(\ar, \ar') \ileadstost (\prop, \text{possible})$, $i$ always considers that $\prop$ is “at least” possible (meaning that $¬\prop$ may be possible as well), when presented with these two arguments. In both cases, the term “always” means that the consideration is stable over time, including after having presented other arguments to $i$. 
Two query protocols presented below will illustrate how these semantics might be satisfied. 
However, I voluntarily leave the querying protocol not precisely specified in general, as multiple reasonable choices are possible. The results of this article depend only on assumptions about $\ileadstost$.

@@

\begin{definition}[Decision situation]
	A decision situation is a topic $\allprops$ closed under negation, a set of arguments $\allargs$, an “always leads to” relation $\ileadstost \subseteq (\allargs × \allargs) × \Phi$ satisfying (A1) and (A2) and using unordered pairs of arguments, $\Phi$ being the set of propositions $\allprops × \set{\text{sure}, \text{possible}}$.
\end{definition}

@@

It is also necessary to capture the evolution of the position of $i$ towards arguments over time: a repeated query using a given pair of arguments may yield different answers, for example because the repeated question has been interleaved with another query containing other arguments (which $i$ may still have in mind when answering the second repetition), or for any other known or unknown reason. The important information for us is the \emph{set} of answers that $i$ could give to a query involving two given arguments and a given topic, among the set $\{\prop, ¬\prop, \text{both}\}$. The set designates all possible answers over time and over different order of asking the queries.
Hence, given $\prop \in \allprops$, define a relation $\ileadsto^\prop \subseteq (\allargs × \allargs) × \powersetz{\{\prop, ¬\prop, \text{both}\}}$, where $\powersetz{B}$ designates the subsets of $B$ excluding the emptyset, with the semantics that $(\ar, \ar') \ileadsto^\prop C$, with $C \subseteq \set{\prop, ¬\prop, \text{both}}$, iff $\forall c \in C$, it is observable at least once that $i$, when presented with $(\ar, \ar')$, considers $c$ valid in her current state of mind. The querying protocol is supposedly made so that for any $\prop \in \allprops$, $\ileadsto^\prop = \ileadsto^{¬\prop}$.

@@

Define $\ileadstost$ from $\ileadsto^\prop$ as follows, for any $\prop \in \allprops$: $(\ar, \ar') \ileadstost (\prop, \text{possible}) ⇔ (\ar, \ar') \ileadsto^\prop C$ with $C \subseteq \set{\prop, \text{both}}$; and $(\ar, \ar') \ileadstost (\prop, \text{sure}) ⇔ (\ar, \ar') \ileadsto^\prop \set{\prop}$.

@@

@@

\section{Empirical theories}
\subsection{Simple to complex}
From an additive single-answer empirical theory, we can define a function: $\forall \ar_1 \gbeats \ar_0: f(\ar_1, \ar_0) = \ar_0 + {\gbeatsinv}(\ar_1) = \set{\ar_0 + \ard_2 \suchthat \ard_2 \in {\gbeatsinv}(\ar_1)}$, and $f(\ar_1, \ar_0) = \emptyset$ otherwise. Thus, $f = \set{((\ar_1, \ar_0), \ar_2) \suchthat \exists \ard_2 \in \allargs \suchthat \ard_2 \gbeats \ar_1 \gbeats \ar_0 \land \ar_2 = \ar_0 + \ard_2}$.
Also, from the previous definition, $g(\ar_0)(\ar_0 + \argsd_2) = dom(f^{-1}(\ar_0 + \argsd_2) \cap (\allargs × \set{\ar_0})) = {\gbeats}(\argsd_2) \cap {\gbeatsinv}(\ar_0)$ because $(\ar_1, \ar_0) \in f^{-1}(\ar_0 + \argsd_2) ⇔ \exists \ar_2 \in \ar_0 + \argsd_2 \suchthat \ar_2 \in f(\ar_1, \ar_0) ⇔ \ar_1 \gbeats \ar_0 \land \exists \ard \in \argsd_2 \suchthat \ard \gbeats \ar_1$.
\commentOC{Could be false.}

@@

Given any $(\allprops, ¬, \allargs, I)$, an \emph{observation protocol} is a binary tree whose non-leaf nodes are among $I × \allhist × \Phi$ and whose two edges out of any non leaf node are labeled by “yes” and “no”; the leaf nodes being simply $\emptyset$. The observation protocol is said to assume memory iff for each non-leaf node $n = (i, \hist, \phi)$ about $i$, any node $n' = (i, {\hist}', \phi')$ about $i$ that is further down the tree continues the history, that is, the sequence ${\hist}'$ starts with $\hist$. Given an observation protocol and observables $(\ileadstoe[i])_{i \in I}$, the \emph{observed path} refers to the path in the tree defined recursively as consisting of the root node and, for each non-leaf node $n = (i, \hist, \phi)$ in the path, the child node along the branch “yes” if $\hist \ileadstoe[i] \phi$ and the child node along the branch “no” otherwise. The \emph{observations} refer to the set of nodes along the observed path.

Given any $(\allprops, ¬, \allargs, I)$, a \emph{normative claim} about $(\allprops, ¬, \allargs, I)$ (or about $\allargs$, when the other defined objects are left implicit) is any first-order logic proposition involving the atoms $[(\hist, \phi) \in {\ileadstoe[i]}], \hist \in \allhist, \phi \in \Phi$ which, together with the axioms, permits to deduce that $\phi \in \iPhi$ for some $i \in I$ and $\phi \in \Phi$. The claim is said to be \emph{about} the given set $\allargs$. An observation protocol, together with observables $(\ileadstoe[i])_{i \in I}$, \emph{falsifies} a normative claim iff the falsity of the claim is deducible from the axioms and the observations. A normative claim is \emph{falsifiable} iff there exists a finite observation protocol such that for any observables $(\ileadstoe[i])_{i \in I} \in (\allhist × \Phi)^I$, if the claim is false given these observables, then it is possible to deduce it from the axioms and the observations given the observables.

@@

\begin{proposition}[Trivial claims are not falsifiable]
	Given any $(\allprops, ¬, \allargs ≠ \emptyset, I)$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Consider any claim about some $i$ and some $\phi$ and any finite observation protocol. We have to show that for some observables, the claim is false, but it is impossible to observe it (meaning, to deduce it from the axioms and observations).

@@

There is another reason for non falsifiability, which does not involve the problem of finiteness.
\begin{proposition}[Minimal non-trivial claims are not falsifiable]
	Given any $(\allprops, ¬, \allargs, I), \card{\allargs} ≥ 2$, no minimal normative claim is falsifiable.
\end{proposition}
\begin{proof}
	Define $\ar_1$ as the first element of the sequence $\hist$ of the root node of the observation protocol (if the observation protocol is empty, suffices to define any observables that make the claim false). Define the observables so that $\forall \hist \in \allhist: \ar_2, \hist \ileadsto ¬\phi$. This proves that no argument is decisive, hence $\phi \notin \Box\Phi$, and the claim is false. But the observables may also be defined so as to satisfy the protocol (defined as: when going through the protocol, the observations can’t fail the claim).
\end{proof}
